---
title: "Lab 3"
format: gfm
editor: source
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
library(knitr)
library(tidyverse)
library(MCMCpack)

```


Consider data from on co-infections of Australian Flying Foxes in the paper [Synchronized seasonal excretion of multiple coronaviruses coincides with high rates of coinfection in immature bats](https://www.nature.com/articles/s41467-025-61521-7). In this paper, we studied co-infections of different coronavirus clades in 176 juvenile flying foxes. Of those,

- 148 bats tested negative for both clades
- 12 tested positive for clade iv but negative for clade v
- 8 tested positive for clade v but negative for clade iv 
- 8 tested positive for both clades



#### 1. (4 points)

Use the `waffle` package to visualize the four outcomes. Add a short caption to your figure. Note you might need to install `waffle` from GitHub.

```{r}
#devtools::install_github("wch/Rttf2pt1")
#devtools::install_github("wch/extrafont")
#devtools::install_github("hrbrmstr/waffle")
library(waffle)

```


#### 2. (2 points)

Create a 2-by-2 contingency table with the results. Note in the analysis, we used the following row / column structure.


```{r}
#| echo: false
tibble(` ` = c('V -', "V +"), `IV -` = c('',''), `IV +` = c('','')) |> kable()
```


#### 3. 

#### 3.1 (2 points)

Use maximum likelihood to calculate point estimates of marginal probabilities of a positive result for each virus.

$Pr[IV(+)]:$ = 

$Pr[V(+)]:$: = 

#### 3.2 (4 points)

Use maximum likelihood to calculate the conditional probabilities for positive result for each virus.

- $Pr[IV (+) |V(+)]$
- $Pr[IV (+) |V(-)]$
- $Pr[V (+) |IV(+)]$
- $Pr[V (+) |IV(-)]$

#### 3.3 (4 points)

Recall that statistical indendence implies that $\pi(A) = \pi(A|B)$. In question, 3.2, we have estimates of $\pi$. Without a formal hypothesis test, do you think that getting a positive test for one clade is related to getting a positive test in the other clade? 


#### 4. 

#### 4.1 (4 points)

Revisit 3.1 and estimate these marginal probabilities with appropriate uncertainty intervals. You are welcome to use MLE or Bayesian approaches.

$Pr[IV(+)]:$ = 

$Pr[V(+)]:$: = 

#### 4.2 (4 points)

Compute an uncertainty interval for the difference in these probabilities in part 4.1.

#### 5 

#### 5.1 (4 points)

Use a $\chi^2$ test to determine whether positivity across the two clades is related.

#### 5.2 (4 points)

Write a paragraph describing the results from 5.1 in the context of the scientific research question.